---
title: "Aides Drug"
author: "Ashley Saunders"
date: "09/24/2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

```{r}
female<- filter(triUmp, gender=="F")
```


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=female)
barchartGC(~infected + treatment,data=female, type="percent")

```
This graph makes it clear that in a study consisting of independent variables, females, those who take the drug are more likely to get aids than those who do not take the drug. 

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
table1 <- xtabs(~infected + treatment,data=female)
rowPerc(table1)
colPerc(table1)
```
The Numerical Descriptive Results clearly show that the percentage of the group that didn't get aids is nearly 41%, while the percentage of the group that got aids is nearly 60%.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
In total, there were three statistical tests ran and with each test we saw the probability of females getting aids when they use this drug or the placebo. The first test showed a probability of 0.00007686, the second 0.0001, and the third 0.00006596. These probabilities show statistical differences between female focused groups and the drug and placebo and whether or not the drug affects results.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
We graphed results of a female focused group and found results of the probability of the differences between getting aids with the drug or the placebo. The bar graph showed that those who used the drug were more likely to get aids than those who took the placebo. Then the numerical graph performed showed that the percent of females who got aids was 20% higher than those who didn't. The three statistical test also showed evidence of females getting aids after using the drug.
